---
title: "GWAS Catalog Data"
author: "jgblanc"
date: "2019-09-30"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)
```

# Data Cleaning  

Filter for:   

  * Study description contains "European"  
  * P-value <= 5.0e-8  
  * Risk allele freq between 0 and 1  


Myopia - pathological myopia - EFO_0004207  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/pathological_myopia.tsv",sep = '\t', header = TRUE)
myopia <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
myopia$disease <- "myopia"
```

Sudden cardiac arrest - EFO_0004278
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/sudden_cardiac_arrest.tsv",sep = '\t', header = TRUE)
sudden_cardiac_arrest <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
sudden_cardiac_arrest$disease <- "sudden_cardiac_arrest"
```

Systemic lupus erythematosus - EFO_0002690
**Check this one** 
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/systemic_lupus_erythematosus.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
systemic_lupus_erythematosus <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
systemic_lupus_erythematosus$disease <- "systemic_lupus_erythematosus"
```

Systemic sclerosis - systemic scleroderma - EFO_0000717  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/systemic_scleroderma.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
systemic_scleroderma <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
systemic_scleroderma$disease <- "systemic_scleroderma"
```

Melonoma - EFO_0000756  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/melanoma.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
melanoma <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
melanoma$disease <- "melanoma"
```

Atrial fibrillation - EFO_0000275  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/atrial_fibrillation.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
atrial_fibrillation <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
atrial_fibrillation$disease <- "atrial_fibrillation"
```

Type II Diabetes - type II diabetes mellitus - EFO_0001360  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/type_2_diabetes.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
type_2_diabetes <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
type_2_diabetes$disease <- "type_2_diabetes"
```

Chronic kidney disease - EFO_0003884  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/chronic_kidney_disease.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
chronic_kidney_disease <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
chronic_kidney_disease$disease <- "chronic_kidney_disease"
```

Age-related macular degeneration - EFO_0001365  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/age_related_macular_degeneration.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
age_related_macular_degeneration <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
age_related_macular_degeneration$disease <- "age_related_macular_degeneration"
```

Primary biliary cholangitis - Primary biliary cirrhosis - EFO_1001486  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/primary_biliary_cholangitis.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
primary_biliary_cholangitis <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
primary_biliary_cholangitis$disease <- "primary_biliary_cholangitis"
```

Vitiligo - EFO_0004208  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/vitiligo.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
vitiligo <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
vitiligo$disease <- "vitiligo"
```

Ankylosing spondylitis - EFO_0003898  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/ankylosing_spondylitis.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
ankylosing_spondylitis <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
ankylosing_spondylitis$disease <- "ankylosing_spondylitis"
```

**Deal with Breast Cancer later** 
Breast cancer - triple negative breast cancer (EFO_0005537), BRACAX breast cancer (EFO_0009443), estrogen-receptor postive breast cancer (EFO_1000649), estrogen-receptor negative breast cancer (EFO_0009781)  
```{r, warning=F, echo=FALSE}
data <- read.csv("data/breast_cancer_all.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data <- subset(data, !grepl("PUBMEDID", data$PUBMEDID))
data$P.VALUE <- as.numeric(as.character(data$P.VALUE))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
breast_cancer <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) 
#%>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
```

Crohn's disease - EFO_0000384  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/crohns_disease.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
crohns_disease <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
crohns_disease$disease <- "crohns_disease"
```

Prostate cancer - prostate carcinoma - EFO_0001663  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/prostate_cancer.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
prostate_cancer <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
prostate_cancer$disease <- "prostate_cancer"
```

Celiac disease - EFO_0001060
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/celiac_disease.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
celiac_disease <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
celiac_disease$disease <- "celiac_disease"
```

Obesity - EFO_0001073  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/obesity.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
obesity <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
obesity$disease <- "obesity"
```

Colorectal cancer - EFO_0005842  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/colorectal_cancer.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
colorectal_cancer <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
colorectal_cancer$disease <- "colorectal_cancer"
```

Rheumatoid Arthritis - EFO_0000685  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/rheumatoid_arthritis.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
rheumatoid_arthritis <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
rheumatoid_arthritis$disease <- "rheumatoid_arthritis"
```

Psoriasis - EFO_0000676  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/psoriasis.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
psoriasis <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
psoriasis$disease <- "psoriasis"
```

Coronary heart disease - EFO_0001645  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/coronary_heart_disease.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
coronary_heart_disease <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
coronary_heart_disease$disease <- "coronary_heart_disease"
```

Testicular germ cell tumor - EFO_1000566  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/testicular_germ_cell_tumor.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
testicular_germ_cell_tumor <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
testicular_germ_cell_tumor$disease <- "testicular_germ_cell_tumor"
```

Type I diabetes - EFO_0001359  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/type_1_diabetes.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
type_1_diabetes <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
type_1_diabetes$disease <- "type_1_diabetes"
```

Ulcerative colitis - EFO_0000729
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/ulcerative_colitis.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
ulcerative_colitis <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
ulcerative_colitis$disease <- "ulcerative_colitis"
```

Schizophrenia - EFO_0000692  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/schizophrenia.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
schizophrenia <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
schizophrenia$disease <- "schizophrenia"
```

Metabolic syndrome - EFO_0000195  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/metabolic_syndrome.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
metabolic_syndrome <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
metabolic_syndrome$disease <- "metabolic_syndrome"
```

Chronic lymphocytic leukemia - EFO_0000095  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/chronic_lymphocytic_leukemia.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
chronic_lymphocytic_leukemia <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
chronic_lymphocytic_leukemia$disease <- "chronic_lymphocytic_leukemia"
```

Multiple sclerosis - EFO_0003885  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/multiple_sclerosis.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
multiple_sclerosis <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
multiple_sclerosis$disease <- "multiple_sclerosis"
```

Asthma - EFO_0000270  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/asthma.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
asthma <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
asthma$disease <- "asthma"
```

Inflamatory bowel disease - EFO_0000270  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/inflamatory_bowel_disease.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
inflamatory_bowel_disease <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
inflamatory_bowel_disease$disease <- "inflamatory_bowel_disease"
```

Parkinson's disease - EFO_0002508  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/parkinsons_disease.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
parkinsons_disease <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
parkinsons_disease$disease <- "parkinsons_disease"
```

Late-onset Alzeheimers - Alzeheimer's disease (late-onset) - EFO_1001870  
```{r, warning=FALSE, echo=FALSE}
data <- read.csv("data/alzheimers_disease.tsv",sep = '\t', header = TRUE, na.strings=c("","NA"))
data$RISK.ALLELE.FREQUENCY <- as.numeric(as.character(data$RISK.ALLELE.FREQUENCY))
alzeheimers_disease <- data %>% select("PUBMEDID", "INITIAL.SAMPLE.SIZE", "RISK.ALLELE.FREQUENCY", "P.VALUE", "SNP_ID_CURRENT", "OR.or.BETA") %>% filter(grepl("European", INITIAL.SAMPLE.SIZE)) %>% filter(P.VALUE <= 5.0e-8) %>% filter(RISK.ALLELE.FREQUENCY < 1 & RISK.ALLELE.FREQUENCY > 0)
alzeheimers_disease$disease <- "alzeheimers_disease"
```


# Re-create Figure 4 Kido et al.  
```{r}
combined <- rbind(age_related_macular_degeneration, alzeheimers_disease, ankylosing_spondylitis, asthma, atrial_fibrillation, celiac_disease, chronic_kidney_disease, chronic_lymphocytic_leukemia,colorectal_cancer, coronary_heart_disease, crohns_disease, inflamatory_bowel_disease, melanoma, metabolic_syndrome, multiple_sclerosis, myopia, obesity, parkinsons_disease, primary_biliary_cholangitis, primary_biliary_cholangitis, prostate_cancer, psoriasis, rheumatoid_arthritis, schizophrenia, sudden_cardiac_arrest, systemic_lupus_erythematosus, systemic_scleroderma, testicular_germ_cell_tumor, type_1_diabetes, type_2_diabetes, ulcerative_colitis, vitiligo)

fig4 <- combined %>% select(RISK.ALLELE.FREQUENCY, disease)
ggplot(data=fig4, aes(x=reorder(disease, RISK.ALLELE.FREQUENCY, FUN = median) , y=RISK.ALLELE.FREQUENCY)) + geom_boxplot(col = "navy", fill = "lightblue") + coord_flip() + ylab("Disease") + xlab("Risk allele frequency") + theme_bw()
```

# Number of Associations per Disease 
```{r}
sample_size <- combined %>% group_by(disease) %>% count()
colnames(sample_size) <- c("disease", "num_associations")
kable(sample_size)
```

# Risk Allele Freq vs Effect Size 
```{r, warning=FALSE, fig.height=8, fig.width=10}
ggplot(data=combined, aes(x = RISK.ALLELE.FREQUENCY, y=OR.or.BETA)) + geom_point(col = "navy") + facet_wrap(~disease, scales = "free") + theme_bw()
```


